Accurate binding of calcium to phospholipid bilayers by effective inclusion of electronic polarization

### 2017-09-29 Manuscript/manuscript.tex
Modifications based on Pavel's comments until the end of section III A.

\begin{figure}[tbp]
  \centering
  \includegraphics[width=7.5cm]{../Fig/POPCstructure.eps}
  \includegraphics[width=8.0cm]{../Fig/ipython_nb/Order-parameters_exp-L14-ECCL17_q80_sig89.eps}
  \includegraphics[height=8.4cm,angle=-90]{../Fig/form-f_exp-l14-eccl17.eps}
  \caption{\label{simVSexpNOions}
    Top: Chemical structure of POPC with definitions of different order parameters calculated.
    Middle: Order parameters of head group, glycerol backbone and sn-1 and sn-2 tails  from simulations with Lipid14 \cite{dickson14} and ECC-lipids models
    compared with experimental order parameters from \cite{ferreira13}.
    Bottom: X-ray scattering form factors from experiments~\cite{Kucerka2011} and simulations using Lipid14 \cite{dickson14} and ECC-lipids models. 
    } 
  \todo{X-label misses glycerol region. JOE: glycerol belongs to the headgroup part, but, yes, it wouldn't hurt to write it explicitly. SAMULI: I would maybe remove the whole text, tell the information in caption and increase the order parameter figure size in z-direction.}
  \todo{Also acyl chain points should correspond the legends and maybe have the same size as HG and glycerol.}
  \todo{Legends would be more clear without multiple points.}
\end{figure}

@@

In order to validate the structural properties of the newly developed ECC-lipids model
we first compared our simulation results without any ions to
NMR order parameters measurements and x-ray scattering form factors 
(Fig.~\ref{simVSexpNOions}). 
As in the original Lipid14 model~\cite{dickson14}, the acyl chain order parameters agree
well with the experimental values. Notably,

and in almost all the other available
lipid models \cite{ollila16}

%This is expected as the ECC model does not modify
%the already highly optimized parameters of this region in Lipid14. 
The headgroup order parameters $\alpha$ and $\beta$ and P-N vector angle are slightly larger
in the ECC-lipid model in comparison to Lipid14 (Figs.~\ref{simVSexpNOions} and \ref{OrderParameterCHANGESsurf}). 
With the current data we cannot, however, conclude which is the more realistic
conformation since in both models one headgroup order parameter agrees
with experiments within error bars whereas the other does not.
Slight changes are also observed in glycerol backbone order parameters,
but the accuracy of both models is comparable to the state of art lipid models
available in literature \cite{ollila16}. The agreement between the x-ray scattering
form factors (Fig.~\ref{simVSexpNOions}) and the areas per molecule from simulations and experiments (Table~\ref{tab:apls}) 
confirm that the membrane structural properties are well captured. 
\todo{JOE: I think the following is actually in Fig.~\ref{simVSexpNOions}.}
A structural comparison of ECC-lipids with Lipid14 can be found 
in SI along with results with other water models. 
%The optimal value for $f_\sigma$ was found by representing the overall membrane structure well
%by matching scattering form factors to experimental data~\cite{Petrache06, Kucerka08, Pabst10}.

@@

\todo{Below analysis is done in a stupid way to get some idea. I would fine it useful to
  do this analysis by using the density profiles, but it is not necessary.}
Rough estimates for the free energy difference between bound and unbound cation
are given by
\begin{equation}
  \Delta G={\rm k_b T}\log(\frac{p_o}{p_i}),
\end{equation}
where $p_o$ and $p_i$ are estimated from the Ca$^{2+}$densities in bulk water and
in the maximum density in bilayer, respectively. The density profiles in
Fig. \ref{fig:cacl-dens} give $\sim 0.8$ k$_{\rm b}$T for the free energy difference
between bound and unbound Ca$^{2+}$ ions in ECC-lipid model and $\sim 1.4$ k$_{\rm b}$T
in Lipid14 with \AA{}qvist.

@@

This section is rough and will likely require editing. 
Data -- up to noted exceptions -- shall be all there, however.

@@

\todo{SAMULI: The same authors have also literature, where they say that ternary complex may not
  be the only option. I will recheck and come back to this.
  SAMULI: This is written in \cite{macdonald87}:
  ''\ce{Ca^{2+}} binding to POPC bilayers over the whole concentration range can be best described in terms
  of formation of a ternary complex involving complexation of two lipids to one calcium ion (Altenbach and Seelig, 1984).
  The addition of a sodium competition term has not changed this conclusion. However, if \ce{Ca^{2+}} concentrations
  up to 100 mM are considered, the data can be equally well explained by a 1:1 binding mechanism (cf. Figure 7).
  In contrast, the Ca2+ binding to POPC-POPG mixtures can be best described by assuming a 1:l stoichiometry regardless of the range of Ca2+concentrations.''
  We might or might not want to discuss about this.
  JOE: I think this is adressed/acknowledged enough now.} \\
\todo{Finalize stoichiometry analysis for \ce{Na^+}, \ce{Ca^{2+}},
  their interaction energies with the lipid membrane, etc,
  and finalize the discussion after these results.}

>>>>

\begin{figure}[tbp]
  \centering
  \includegraphics[width=7.5cm]{../Fig/POPCstructure.eps}
  \includegraphics[width=8.0cm]{../Fig/ipython_nb/Order-parameters_exp-L14-ECCL17_q80_sig89.eps}
  \includegraphics[height=8.4cm,angle=-90]{../Fig/form-f_exp-l14-eccl17.eps}
  \caption{\label{simVSexpNOions}
    Top: Chemical structure of POPC with definitions of different order parameters calculated.
    Middle: Order parameters of head group, glycerol backbone and sn-1 and sn-2 tails  from simulations with Lipid14 \cite{dickson14} and ECC-lipids models
    compared with experimental order parameters from \cite{ferreira13}.
    Bottom: X-ray scattering form factors from experiments~\cite{Kucerka2011} and simulations using Lipid14 \cite{dickson14} and ECC-lipids models. 
    } 
  \todo{X-label misses glycerol region. JOE: glycerol belongs to the headgroup part, but, yes, it wouldn't hurt to write it explicitly. SAMULI: I would maybe remove the whole text, tell the information in caption and increase the order parameter figure size in z-direction.} \\
  \todo{Also acyl chain points should correspond the legends and maybe have the same size as HG and glycerol.} \\
  \todo{Legends would be more clear without multiple points.}
\end{figure}

@@

%In order to validate the structural properties of the newly developed ECC-lipids model
%we first compared our simulation results without any ions to
%NMR order parameters measurements and x-ray scattering form factors 
%(Fig.~\ref{simVSexpNOions}). 
The x-ray scattering form factor (Fig.~\ref{simVSexpNOions}) and
the area per lipid (Table~\ref{tab:apls}) of POPC bilayer simulated
with the ECC-lipid model are in good agreement with the experimental results,
as well as with the Lipid14 model. Thus, we conclude that the lipid bilayer dimensions
are well captured by the developed ECC-lipid model. 
%The optimal value for $f_\sigma$ was found by representing the overall membrane structure well
%by matching scattering form factors to experimental data~\cite{Petrache06, Kucerka08, Pabst10}.

As in the original Lipid14 model~\cite{dickson14}, the acyl chain order parameters of the
ECC-lipid model agree well with the experimental values in Fig.~\ref{simVSexpNOions}. Notably, the experimentally measured forking and
small order parameter values of $C_2$ segment in {\it sn}-2 chain are relatively well
reproduced by the both models. This has been suggested to indicate that the carbonyl
region of {\it sn}-2 chain is directed towards the water phase, in contrast to the
carbonyl in {\it sn}-1 chain, which would orient more along the bilayer
plane~\cite{seelig75,schindler75,gawrisch92}. While this may be an important
feature for the ion binding details, it is not necessarily reproduced by the
available lipid models~\cite{ollila16}
%This is expected as the ECC model does not modify
%the already highly optimized parameters of this region in Lipid14.

The headgroup order parameters of $\alpha$ and $\beta$ carbons are slightly larger
in the ECC-lipid model than in Lipid14, which is apparently related to the
smaller P-N vector angle with respect to the membrane normal in
Fig.~\ref{OrderParameterCHANGESsurf}. With the current data we cannot,
however, conclude which one of the models give the more realistic
headgroup conformations. The ECC-lipid model gives
the $\beta$ carbon order parameter value closer to experiments, while
value for $\alpha$ carbon is better in Lipid14.
Despite of some deviations from the experimental order parameter values
in Fig. \ref{simVSexpNOions},
the accuracy of the both models in the glycerol backbone region
is comparable to the other state
of art lipid models available in literature \cite{botan15}.

@@

\todo{Below analysis is done in a stupid way to get some idea. I would fine it useful to
  do this analysis by using the density profiles, but it is not necessary.
  The reason why I want to do something like this, is that it would be useful to have
  some easier way to estimate the correct binding affinity than the electrometer concept,
  which is quite tedious to apply in practise (simulations with different concenrations,
  cationic surfactant check, etc.). I wuold like to be able to estimate much faster from a
  simulation if the binding affinity is reasonable or not. This may not be the correct way for
that anyway.}
Rough estimates for the free energy difference between bound and unbound cation
are given by
\begin{equation}
  \Delta G={\rm k_b T}\log(\frac{p_o}{p_i}),
\end{equation}
where $p_o$ and $p_i$ are estimated from the Ca$^{2+}$densities in bulk water and
in the maximum density in bilayer, respectively. The density profiles in
Fig. \ref{fig:cacl-dens} give $\sim 0.8$ k$_{\rm b}$T for the free energy difference
between bound and unbound Ca$^{2+}$ ions in ECC-lipid model and $\sim 1.4$ k$_{\rm b}$T
in Lipid14 with \AA{}qvist.

@@

This section is rough and will likely require editing. 
Data -- up to noted exceptions -- shall be all there, however. 
\todo{SAMULI: I think the we should make clear that the current binding
  constants in, e.g., Marsh's handbook are based on the headgroup order parameter
  changes intepreted with ternary complex binding model. I.e. the experimental raw data
  is excatly the same as we have in this work. Our simulations enable much more versatile interpretation
  of the binding phenomena. I think that this is one the main points of this work, so this comment
  applies to the previous section, introduction, abstract and conclusions as well.
}

@@

\todo{Finalize stoichiometry analysis for \ce{Na^+}, \ce{Ca^{2+}},
  their interaction energies with the lipid membrane, etc,
  and finalize the discussion after these results.}
